---
title: "P8451 HW0"
author: "Will Simmons"
date: "1/17/2020"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup}

library(tidyverse)
library(readr)

```

0. Importing data
```{r import}

data = 
  read_csv('./data/dataR2.csv')

```

1. Construct a table providing summaries of the quantitative features of the dataset. Summaries should include the mean, median, minimum value, and maximum value. If you are unable to construct a formatted table within R, you can print raw output, but then comment the output to identify the answer that was requested.

```{r 1_table}

data %>%
  summarize_all(
    list(
      Mean = mean,
      Median = median,
      Minimum = min,
      Maximum = max)
  ) %>% 
  gather(stat, val) %>% 
  separate(stat, into = c("Variable", "stat"), sep = "_") %>% 
  spread(stat, val) %>% 
  select(Variable, Mean, Median, Minimum, Maximum) %>% 
  knitr::kable(
    caption = "Summary of dataset variables"
  )

```

2. Recode BMI into the WHO-defined categories below:
  + Severely underweight - BMI less than 16.5kg/m^2 
  + Underweight - BMI under 18.5 kg/m^2 
  + Normal weight - BMI greater than or equal to 18.5 to 24.9 kg/m^2 
  + Overweight – BMI greater than or equal to 25 to 29.9 kg/m^2 
  + Obesity class I – BMI 30 to 34.9 kg/m^2 
  + Obesity class II – BMI 35 to 39.9 kg/m^2 
  + Obesity class III – BMI greater than or equal to 40 kg/m^2
  
```{r 2_recode_bmi}

data =
  data %>% 
  mutate(
    BMI_new = 
      case_when(
        BMI < 16.5 ~ 'Severely Underweight',
        between(BMI, 16.5, 18.4) ~ 'Underweight',
        between(BMI, 18.5, 24.9) ~ 'Normal Weight',
        between(BMI, 25, 29.9) ~ 'Overweight',
        between(BMI, 30, 34.9) ~ 'Obesity Class I',
        between(BMI, 35, 39.9) ~ 'Obesity Class II',
        BMI >= 40 ~ 'Obesity Class III'
      )
  )

```
  
3. Create a bar chart showing the proportion of breast cancer cases and controls within each BMI category

```{r 3_bar}

data %>% 
  mutate(
    Classification = as.factor(Classification),
    BMI_new = fct_reorder(BMI_new, BMI)
  ) %>% 
  group_by(BMI_new, Classification) %>% 
  summarize(count = n()) %>% 
  mutate(percent = count/sum(count)) %>% 
  
  ggplot(
    aes(
      fill = Classification,
      x = BMI_new,
      y = percent
    )
  ) +
  geom_bar(
    position = 'stack', 
    stat = 'identity'
  ) +
  labs(
    y = 'Proportion\n',
    x = 'BMI Category',
    title = 'Proportion of cases and controls by BMI category'
  ) +
  theme(
    legend.position = 'bottom',
    plot.title = element_text(hjust = 0.5),
    legend.title = element_blank(),
    
  ) +
  scale_fill_discrete(
    labels = c('Healthy controls', 'Breast cancer patients')
  )

```

4. Construct a logistic regression model using breast cancer classification as the outcome and glucose, HOMA, leptin, BMI (continuous) and age as the independent variables. 
  + Fill in the beta estimate and 95% confidence interval associated with a 1-unit change in HOMA

```{r 4_logistic}

logistic_model =
  data %>% 
  mutate(
    Classification = as.factor(Classification),
    Classification = fct_recode(
      Classification,
      '0' = '1',
      '1' = '2'
    )
  ) %>% 
  glm(
    Classification ~ Glucose + HOMA + Leptin + BMI + Age,
    data = .,
    family = binomial
  ) 

logistic_model_tidy =
  logistic_model %>% 
  broom::tidy()

ci_betas =
  as.data.frame(
    confint.default(logistic_model)
  ) %>% 
  add_column(
    Variable = c(
      '(Intercept)',
      'Glucose',     
      'HOMA',  
      'Leptin',      
      'BMI',         
      'Age')
  )

final_table_logistic = 
  inner_join(
    logistic_model_tidy,
    ci_betas,
    by = c('term' = 'Variable')
  ) %>% 
  select(term, estimate, '2.5 %', '97.5 %', everything())

final_table_logistic %>% 
  knitr::kable()

```

The beta estimate for a 1-unit change in HOMA is `r final_table_logistic$estimate[3]`. The associated 95% CI is (`r final_table_logistic$'2.5 %'[3]`, `r final_table_logistic$'97.5 %'[3]`).

5.	Construct a linear regression model using insulin as the outcome and BMI (continuous), age, and glucose as the independent variables. 
  + Fill in the beta estimate and 95% confidence interval associated with a 1-unit change in age.
  
```{r 5_linear}

linear_model =
  data %>% 
  lm(
    Insulin ~ BMI + Age + Glucose,
    data = .
  ) 

linear_model_tidy =
  linear_model %>% 
  broom::tidy()

ci_betas_linear =
  as.data.frame(
    confint.default(linear_model)
  ) %>% 
  add_column(
    Variable = c(
      '(Intercept)',
      'BMI',     
      'Age',  
      'Glucose')
  )

final_table_linear = 
  inner_join(
    linear_model_tidy,
    ci_betas_linear,
    by = c('term' = 'Variable')
  ) %>% 
  select(term, estimate, '2.5 %', '97.5 %', everything())

final_table_linear %>% 
  knitr::kable()

```

The beta estimate for a 1-unit change in age is `r final_table_linear$estimate[3]`. The associated 95% CI is (`r final_table_linear$'2.5 %'[3]`, `r final_table_linear$'97.5 %'[3]`).
